---
title: "An R package for the ARBP"
output: github_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "plots/README-"
)
```

The `ramboseli` package aims to provide a set of small functions and utilities for R that can be shared among members of the Amboseli Baboon Research Project. Currently it's just a barebones proof-of-concept, but this documentation will grow as features are added to the package.

## Preparation

To use this package, you first need to install the `devtools` package:

```{r, eval = FALSE}
    install.packages("devtools")
```

Then, you can install the latest development version of `ramboseli` from github:

```{r, eval = FALSE}
  devtools::install_github("amboseli/ramboseli")
```

After you have installed the package once, you can simply load it in the future using:

```{r, message = FALSE, warning = FALSE} 
  library(ramboseli)
```


## Current functionality provided by the package:

- [Plotting helpers and functions](documentation/plotting.md)


## Guidelines for using `ramboseli` functions

TO DO: Expand this section

Instructions to be added:
* Detailed instructions for linking babase to R via ssh tunnel
* Managing passwords and keeping them of your scripts: ssh key or rstudioapi::askForPassword()
* Creating connection to database using DBI::dbConnect()
* Best practices for supplying babase tables as arguments to functions
  1. Supply a database connection and no tables to obtain data directly through queries
    - Uses most up-to-date data
    - Use when carrying out exploratory analysis
    - Not reproducible
  2. Supply tables from the R environment and no database connection
    - Uses static input files (a "snapshot" of specific babase tables)
    - Reqires that user queries database first to produce static files and loads them into R environment
    - Use for final analysis if you desire reproducibility
